---
output: github_document
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# eqvis

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/akbarpunjwani/msdr-capstone-eqvis.svg?branch=main)](https://travis-ci.org/akbarpunjwani/msdr-capstone-eqvis)
<!-- badges: end -->

The goal of EQVIS is to provide EarthQuake VISualization tools in order to analyze the earthquakes occurrences as reported at NOAA sample data set as available for public access.

## Installation

You can install the released version of eqvis from [CRAN](https://CRAN.R-project.org) with:

``` r
library(devtools)
install_github("akbarpunjwani/msdr-capstone-eqvis")
library(eqvis)
```

The EQVIS, which stands for EarthQuake VISualizations, provide a sample data set for public access to analyze the earthquakes occurrences as reported at [NOAA](https://www.ngdc.noaa.gov/nndc/struts/form?t=101650&s=1&d=1). 

This package includes this data for internal use only. However, this package uses that data and provides a **geometric objects** (*eq_timeline_and_labels*) and the **mapping based visual plots** (*eq_map_with_datelabels and eq_map_with_htmllabels*) for specified year.

## Data Summary ##
The NOAA Earthquakes data has a time series information of DAY, MONTH and YEAR of the earthquakes along with the spatial information of Longitude, Latitude & Location. Additionally, it also has the data of eq magnitude & total deaths. In order to load & clean data, below methods can be used:

**eq_load_data(filename)**
Since the package includes the dataset of NOAA for earthquakes, the output would be the raw data as loaded from the archived file available with public access.

Examples:
*eqvis::eq_load_data()*
```{r}
eqvis::eq_load_data()
```

*eqvis::eq_location_clean()*
```{r}
eqvis::eq_location_clean()
```

*eqvis::eq_clean_data()* (Internally loads raw data when no data is passed)
```{r echo = TRUE, eval=TRUE, warning=TRUE}
eqvis::eq_clean_data()
```


```{r  messages = FALSE, echo = FALSE, eval=TRUE, warnings=TRUE, include=TRUE, results='hide', fig.height=2, fig.width=2}
sample <- 1:10
```

## Data Visualization ##
The NOAA Earthquake data contains details (DATE, MAGNITUDE, TOTAL DEATHS, LOCATION) of world-wide occurrences of earthquakes. The package uses this data to visualize it into a timeline plot for each country, using the below mentioned method:

**eqvis::eq_timeline()**

As the package includes the dataset of NOAA earthquakes, this method helps the user to visualize the data for the selected countries, time series range & magnitude.
In case any of the required details not provided, it provides the sample output for specific values.

Examples:
*eqvis::eq_timeline(countries = c('TURKEY','CHINA'))*

```{r fig.height=5, fig.width=7}
grid::grid.draw(eqvis::eq_timeline(countries = c('TURKEY','CHINA')))
```

**eqvis::eq_timeline_and_labels()**

This method helps the user to visualize the largest earthquakes locations in addition to the time series range & magnitude.
In case any of the parameters is not provided with values, it displays the sample with example values.

Examples:
*eqvis::eq_timeline_and_labels(countries = c('TURKEY','CHINA'))*

```{r fig.height=5, fig.width=7}
grid::grid.draw(eqvis::eq_timeline_and_labels(countries = c('TURKEY','CHINA')))
```

**eqvis::eq_map_with_datelabels()**

As the package includes the dataset of NOAA earthquakes, this method helps the user to get a interactive map of earthquakes along with popup text showing date labels.

Examples:
*eqvis::eq_map_with_datelabels()*

```{r fig.height=5, fig.width=7}
eqvis::eq_map_with_datelabels()
```

**eqvis::eq_map_with_htmllabels()**

When compared with eq_map_with_datelabels, this function shows the popup text having more details and formatted using HTML tags. Instead of date, the popup text shows Clean Location, Magnitude & Total Deaths.

Examples:
*eqvis::eq_map_with_htmllabels()*

```{r fig.height=5, fig.width=7}
eqvis::eq_map_with_htmllabels()
```
