---
title: "Earthquake Visualization Tools"
description: "A meaningful explaination of all the Geometric tools & Mapping Leaflet Markers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Earthquake Visualization Tools}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eqvis)
```
```{r  global_options}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,
  warning = FALSE)
```

This tutorial is about EQVIS package and how it could be used.

The EQVIS, which stands for EarthQuake VISualizations, provid a sample data set for public access to analyze the earthquakes occurrences as reported at NOAA. 

This package includes this data for internal use only. However, this package uses that data and provides a **geometric objects** (*eq_timeline_and_labels*) and the **mapping based visual plots** (*eq_map_with_datelabels and eq_map_with_htmllabels*) for specified year.

## Data Summary ##
The NOAA Earthquakes data has a time series information of DAY, MONTH and YEAR of the earthquakes along with the spatial information of Longitude, Latitude & Location. Additionally, it also has the data of eq magnitude & total deaths. In order to load & clean data, below methods can be used:

**eq_load_data(filename)**
Since the package includes the dataset of NOAA for earthquakes, the output would be the raw data as loaded from the archived file available with public access.

Examples:
*eqvis::eq_load_data()*
```{r}
eqvis::eq_load_data()
```

*eqvis::eq_location_clean()*
```{r}
eqvis::eq_location_clean()
```

*eqvis::eq_clean_data()* (Internally loads raw data when no data is passed)
```{r echo = TRUE, eval=TRUE, warning=TRUE}
eqvis::eq_clean_data()
```


```{r  messages = FALSE, echo = FALSE, eval=TRUE, warnings=TRUE, include=TRUE, results='hide', fig.height=2, fig.width=2}
sample <- 1:10
```

## Data Visualization ##
The NOAA Earthquake data contains details (DATE, MAGNITUDE, TOTAL DEATHS, LOCATION) of world-wide occurrences of earthquakes. The package uses this data to visualize it into a timeline plot for each country, using the below mentioned method:

**eqvis::eq_timeline()**

As the package includes the dataset of NOAA earthquakes, this method helps the user to visualize the data for the selected countries, time series range & magnitude.
In case any of the required details not provided, it provides the sample output for specific values.

Examples:
*eqvis::eq_timeline(countries = c('TURKEY','CHINA'))*

```{r fig.height=5, fig.width=7}
grid::grid.draw(eqvis::eq_timeline(countries = c('TURKEY','CHINA')))
```

**eqvis::eq_timeline_and_labels()**

This method helps the user to visualize the largest earthquakes locations in addition to the time series range & magnitude.
In case any of the parameters is not provided with values, it displays the sample with example values.

Examples:
*eqvis::eq_timeline_and_labels(countries = c('TURKEY','CHINA'))*

```{r fig.height=5, fig.width=7}
grid::grid.draw(eqvis::eq_timeline_and_labels(countries = c('TURKEY','CHINA')))
```

**eqvis::eq_map_with_datelabels()**

As the package includes the dataset of NOAA earthquakes, this method helps the user to get a interactive map of earthquakes along with popup text showing date labels.

Examples:
*eqvis::eq_map_with_datelabels()*

```{r fig.height=5, fig.width=7}
eqvis::eq_map_with_datelabels()
```

**eqvis::eq_map_with_htmllabels()**

When compared with eq_map_with_datelabels, this function shows the popup text having more details and formatted using HTML tags. Instead of date, the popup text shows Clean Location, Magnitude & Total Deaths.

Examples:
*eqvis::eq_map_with_htmllabels()*

```{r fig.height=5, fig.width=7}
eqvis::eq_map_with_htmllabels()
```
